\begin{thebibliography}{1}
\selectlanguage{english}

% links de partikle
\bibitem[1]{partikleos}
	S. Peiro, M. Masmano, I. Ripoll, and A. Crespo,
	\newblock { \em{PaRTiKle OS, a replacement of the RTLinux core}}.
	\newblock Real-Time Systems Group, Universidad Politecnica de Valencia.
	\hrefx{http://www.e-rtl.org/partikle/node/5},
	\hrefx{http://www.e-rtl.org/partikle}.
	
\bibitem[2]{partikle-um}
	M. Masmano, I. Ripoll, A. Crespo,
	\newblock { \em{PaRTiKle OS User Manual}}.
	\newblock Real-Time Systems Group, Universidad Politecnica de Valencia.
	\hrefx{http://www.e-rtl.org/partikle/node/5}.
	
\bibitem[3]{xtratum}
	M. Masmano; I. Ripoll, A. Crespo, 
	\newblock {\em {An overview of the XtratuM nanokernel}}, 
	\newblock  Workshop on Operating Systems Platforms for Embedded Real-Time applications, (2005),
	\hrefx{http://www.xtratum.org/biblio}.

% links de LPC
\bibitem[3]{lpc213x-um}
	NXP,
	\newblock { \em {LPC2131/2/4/6/8 User manual}}.
	\newblock NXP, founded by Philips,
	\newblock \hrefx{http://www.standardics.nxp.com/support/documents/microcontrollers/?search=LPC2\&type=user}.

\bibitem[4]{tigtarm7}
	Trevor Martin,
	\newblock { \em {The Insider's guide to the ARM7-based microcontrollers}}.
	\newblock Hitex development tools,
	\hrefx{http://hitex.co.uk/arm}.

\bibitem[5]{arm7tdmi-s}
	Advanced RISC Machines,
	\newblock { \em{ARM7TDMI-S (rev r4p3) Technical Reference manual}}.
	\newblock ARM Limited.
	\hrefx{http://www.arm.com/documentation/ARMProcessor\_Cores}.

\bibitem[6]{armarm}
	David Seal,
	\newblock { \em{The ARM Architecture Reference Manual}}.
	\newblock 2nd Edition, Addison-Wesley Longman Publishing Co.
	\hrefx{http://www.arm.com/documentation/books.html}.

\bibitem[7]{lpc2000ans}
	NXP,
	\newblock { \em{LPC2000 Application Notes, Boot sequence, Interrupts, Spurious Interrupts}}.
	\newblock NXP, founded by Philips.
	\hrefx{http://www.standardics.nxp.com/support/documents/microcontrollers/?search=LPC2}.

\bibitem[8]{lpc2k_pgm}
	Paul Stoffregen,
	\newblock { \em{LPC2K\_PGM Linux Bootloader Utility (Philips LPC 2000 ARM7 Chips)}}.
	\hrefx{http://www.pjrc.com/arm/lpc2k\_pgm}.

% misc links
\bibitem[9]{mikro}
	\newblock { \em{The Mikrocontroller project}},
	\hrefx{http://www.mikrocontroller.net/articles/LPC2000}.
	
\bibitem[10]{lpc-examples}
	Martin Thomas,
	\newblock { \em{Philips LPC213x/214x examples ported for the GNU-Toolchain}},
	\hrefx{http://www.siwawi.arubi.uni-kl.de/avr\_projects/arm\_projects/lpc2k\_bundle\_port/}.

\bibitem[11]{ihex}
	Intel Corporation,
	\newblock { \em{Intel Hexadecimal Object File Format Specification}},
	\newblock  Intel 1988, 
	\hrefx{http://www.pjrc.com/tech/8051/pm2\_docs/intel-hex.html}.

	
\bibitem[12]{FTDI}
	Craig Peacock,
	\newblock {\em USB with the simplicity of RS-232}
	\hrefx{http://www.beyondlogic.org/usb/ftdi.htm}.

\bibitem[13]{UART}
	Craig Peacock,
	\newblock {\em Interfacing the Serial / RS-232 Port}
	\hrefx{http://www.beyondlogic.org/serial/serial.htm}

\bibitem[14]{ATCPS}
	Development systems Division, Compiler Tools Group
	\newblock {\em Procedure Call Standard for the ARM Architecture}.
	\newblock  ARM Limited, 2003.
	\hrefx{http://infocenter.arm.com/help/topic/com.arm.doc.ihi0042a/IHI0042A\_aapcs.pdf}
	
\bibitem[15]{lamport}
            Leslie Lamport.
            \newblock {\em {\LaTeX:} {A} Document Preparation System.
            (User's Guide and Reference Manual)}.
            \newblock Addison-Wesley, Reading, Massachusetts, second edition, 1986.
\end{thebibliography}
